---
title: "R Fundamentals"
author: "Prof. Andre Salerno, MSc"
date: '2019-04-14'
categories: ["R"]
tags: ["fundamentals", "commands"]
draft: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, warning = FALSE)
load("C:/Users/andre/r_programming/blogdown/env.RData")
```

It is very important in terms of computing knowledge that during your daily jorney with the R Programming you should make a notebook with the principals commands, like a computing R library. Because of that I have decided done this alphabetical list with the major commands in R. Enjoy it!

### Asign

```{r}
z <- x + y
z
```


### Basic operations

```{r}
n + 5
```

### Code Flow

```{r}
if (x < y){
  print("x is lesser than y")
} else if (x == y){
  print("x is equal to y")
} else {
  print("x is greater than y")
}

for (i in 1:10) {
  print(i)
}

t <- matrix(1:6, nrow = 2, ncol = 3)
t

for (i in seq(nrow(t))) {
  for(j in seq(ncol(t))){
    print(t[i,j])
  }
}
```


### Data Frames

```{r}
o <- c(1, 3, 5)
p <- c("red", "yellow", "blue")
q <- c(TRUE, FALSE, TRUE)
df <- data.frame(o,p,q)
df

df[2,1]
df[,"p"]

r <- iris

ncol(r)

nrow(r)

head(r, 2)

str(r)
```


### Division of variables

```{r}
x / y
```

### Exponential

```{r}
exp(y)
```

### Factor

```{r}
f <- c(1,4,7,9,8,1)
f <- factor(f)
f
levels(f)

factor(f, exclude = 4)

g <- c(1,2,7,7,1,2,2,7,1,7)
g <- factor(g, levels = c(1,2,7), ordered = TRUE)
g
```


### Factorial

```{r}
factorial(y)
```

### Functions

```{r}
squaredNum <- function(number){
  s <- number^2
  return(s)
}

squaredNum(25)
```


### Lists

```{r}
m <- list(1:4, "book", 1+4i)
m

m[[1]][2] # accessing elements in a list
```


### Logical operators

```{r}
x > y

x + y != 8

x == y

x & y > 2

isTRUE(x < y)
```

### Matrices

```{r}
i <- matrix(c(1,2,3,4,5,6), nrow = 2, ncol = 3, byrow = TRUE, dimnames = list(c("Obs1", "Obs2"), c("Col1", "Col2", "Col3")))
i

j <- rbind(1:3, 4:6, 10:12)
j

k <- cbind(1:3, 4:6, 10:12)
k

l <- array(c(1,2,3,4,5,6), dim = c(2,3))
l

l[1,2]
```

### Matrix Functions

```{r}
diag(l) # matrix diagonal

dim(l) # dimensions

colSums(l) # column sums

rowSums(l) # row sums

t(l) # matrix transpose

```


### Maximum value

```{r}
max(x,y)
```

### Minimum value

```{r}
min(x,y)
```


### Multiplication of variables

```{r}
x * y
```



### Objects

```{r}
n <- 10
```

### pi number:

```{r}
x * pi
```


### Power of variables

```{r}
x ** y
```

### Sign of a number

```{r}
sign(y)
```

### Square root

```{r}
sqrt(y)
```

### String

```{r}
h <- "This is our first string"
class(h)
```

### String function

```{r}
nchar(h)

substring(h, 4)

substring(h, 1,4)

strsplit(h, " ")

grep("our", h)

grep("book", h)

regexpr("our",h)

tolower(h)

toupper(h)

sub("our", "my", h)

paste(h, "My book", sep = " : ")
```



### Subtractions of variables

```{r}
x - y
```


### Sum variables

```{r}
x <- 4
y <- 3

x + y
```

## Vector

```{r}
a <- c(1,3,5,8)
class(a)

a + 3 ## adding 3

b <- c(1:10) # consecutive numbers
b

c <- seq(1, 5, by = 0.5)
c

d <- seq(1, 5, length.out = 22)
d

e <- rep(22, 4)
e

rev(b) # reverse
```

### Vector indexing

```{r}
a[-3] # excluding third element

a[2] # choosing second element

a[1:2] # choosing sequence

a[a>2 & a<6]

a[c(TRUE, FALSE, FALSE, TRUE)]
```

### Vector functions

```{r}
sum(a)

prod(a)

length(a)

a[1] <- 4
a
```

